Who does big team science?

### 2023-08-16 manuscript_scopus.Rmd
adding the discussion

This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (https://osf.io/f2dtr) but then was updated with access to the Scopus database for a broader picture of BTS projects (https://osf.io/fheun). All materials and code can be found on our OSF page: https://osf.io/cgx6u/ or corresponding GitHub archive: https://github.com/doomlab/big_team_who.

@@

### RQ1: Publisher Information.

@@

### RQ2: Publication Information.

@@

### RQ3: Author Information.

@@

Figure \@ref(fig:fig-career) portrays the average career length for authors involved in BTS publications across years. Career length was defined as the year of first publication minus the current year, and higher numbers mean longer careers. To analyze trends over time, we  calculated the average career length for each publication (i.e., average author career lengths to create one score for each paper) and analyzed a regression analysis using career length to predict year of publication. In order to show variance between individuals, we calculated the standard deviation of career length for each publication and used variance as an additional predictor.

@@

We used the same analyses described in the career length section to analyze trends over time. An increasing slope over time indicates that individuals who are publishing more are more represented in BTS over time (i.e., increasing numbers of scholars with higher publication rates), while a negative slope indicates more researchers with less publications. A positive slope for the standard deviation of publication metrics indicates increasing variance over time (i.e., more diversity in the individual publication rates), while a negative slope would indicate less diversity in researchers over time. While publication rates do not represent value as a researcher, they are often used in hiring and promotion decisions, and we used this variable as a proxy to gauge the diversity in scholars represented in big teams. As shown in Figure \@ref(fig:fig-heatmap) publication metrics were generally negative for the average publication metrics, indicating more scholars over time with lower numbers of citations with the strongest effects in health and social sciences. The variability of publication counts was not significant for the life sciences but was negative for the physical sciences (less variability over time) and positive for social and health sciences (more variability and over time). This result indicates that the physical sciences are trending toward scholars with less publications but also less diverse in number of publications, while the health and social sciences see more diversity in publication counts and less published scholars overall.

@@

Author geopoligical region is displayed in Figure \@ref(fig:fig-map-both). Big team publications appear to be lead by North America and Western Europe, while all publications are lead by North America and East Asia. To understand the change in representation diversity, we examined if the number of regions in a publication is predicted by the year of publication. Increasing diversity would be represented by a positive slope, while decreasing diversity would be represented by a negative slope. As shown in Figure \@ref(fig:fig-heatmap), the physical sciences do not show a trend of change in representation, while all other sciences showed a positive effect increasing in the number of geopolitical regions authors represent on publications.

@@

Figure \@ref(fig:fig-author-gpe) indicates the percent of authors in regions. In general, we found the same pattern as the overall analysis wherein most authors are from Europe and North America. The pattern of representation is roughly similar for the separation of small, medium, and large numbers of authors on papers. Across time, the representation does appear to diversify, with more representation in Asia, Latin American, and Africa. Figure \@ref(fig:fig-effect-gpe) represents the size of the differences in first/corresponding authors and other authors across time and number of authors. The differences in representation are larger for papers with more authors; however, the effects are non-zero for many of the comparisons. Encouragingly, over time these effects appear to diminish in size. One limitation with the calculation of effect sizes for count data is the sensitivity of the data to sample size (i.e., $\chi^2$ is upwardly biased by sample size, and $V$ is calculated based on this value). While we used the inclusion of zero as our boundary for "significance", the interpretation of the effects is that most are likely small: $V$ < .05: `r apa_num(sum(all_effects$v <.05)/nrow(all_effects)*100)`%, < .10: `r apa_num(sum(all_effects$v < .10)/nrow(all_effects)*100)`%, < .20: `r apa_num(sum(all_effects$v < .20)/nrow(all_effects)*100)`%.

@@

- number of publications increasing
- research areas appear to be cancer, physics, and psychology
- the number of authors is increasing as well 
- career length is decreasing, number of publications necessary to be involved is decreasing
- number of gpe increasing 
- appears to be slowly diversifying, yet still not equal in first and other authors in diversity

limitations - based on the data we could get, we made up the definition of BTS, under-representation of articles in other languages/that they don't collect

future: incentives for bts and why people do it

@@

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```

::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\endgroup
```

# (APPENDIX) Appendix {-}

>>>>

This manuscript was preregistered with the same conceptual ideas using
Google Scholar and ORC-ID databases (<https://osf.io/f2dtr>) but then
was updated with access to the Scopus database for a broader picture of
BTS projects (<https://osf.io/fheun>). All materials and code can be
found on our OSF page: <https://osf.io/cgx6u/> or corresponding GitHub
archive: <https://github.com/doomlab/big_team_who>.

@@

### RQ1: Publisher Information

@@

### RQ2: Publication Information

@@

### RQ3: Author Information

@@

Figure \@ref(fig:fig-career) portrays the average career length for
authors involved in BTS publications across years. Career length was
defined as the year of first publication minus the current year, and
higher numbers mean longer careers. To analyze trends over time, we
calculated the average career length for each publication (i.e., average
author career lengths to create one score for each paper) and analyzed a
regression analysis using career length to predict year of publication.
In order to show variance between individuals, we calculated the
standard deviation of career length for each publication and used this
variance as an additional predictor.

@@

We used the same analyses described in the career length section to
analyze trends over time. An increasing slope over time indicates that
individuals who are publishing more are more represented in BTS over
time (i.e., increasing numbers of scholars with higher publication
rates), while a negative slope indicates more researchers with less
publications. A positive slope for the standard deviation of publication
metrics indicates increasing variance over time (i.e., more diversity in
the individual publication rates), while a negative slope would indicate
less diversity in researchers over time. While publication rates do not
represent value as a researcher, they are often used in hiring and
promotion decisions, and we used this variable as a proxy to gauge the
diversity in scholars represented in big teams. As shown in Figure
\@ref(fig:fig-heatmap) publication metrics were generally negative for
the average publication metrics, indicating more scholars over time with
lower numbers of publications with the strongest effects in health and
social sciences. The variability of publication counts was not
significant for the life sciences but was negative for the physical
sciences (less variability over time) and positive for social and health
sciences (more variability and over time). This result indicates that
the physical sciences are trending toward scholars with less
publications but also less diverse in number of publications, while the
health and social sciences see more diversity in publication counts and
less published scholars overall.

@@

Author geopolitical region is displayed in Figure
\@ref(fig:fig-map-both). Big team publications appear to be lead by
North America and Western Europe, while all publications are lead by
North America and East Asia. To understand the change in representation
diversity, we examined if the number of regions in a publication is
predicted by the year of publication. Increasing diversity would be
represented by a positive slope, while decreasing diversity would be
represented by a negative slope. As shown in Figure
\@ref(fig:fig-heatmap), the physical sciences do not show a trend of
change in representation, while all other sciences showed a positive
effect increasing in the number of geopolitical regions authors
represent on publications.

@@

Figure \@ref(fig:fig-author-gpe) indicates the percent of authors in
regions. In general, we found the same pattern as the overall analysis
wherein most authors are from Europe and North America. The pattern of
representation is roughly similar for the separation of small, medium,
and large numbers of authors on papers. Across time, the representation
does appear to diversify, with more representation in Asia, Latin
American, and Africa. Figure \@ref(fig:fig-effect-gpe) represents the
size of the differences in first/corresponding authors and other authors
across time and number of authors. The differences in representation are
larger for papers with more authors; however, the effects are non-zero
for many of the comparisons. Encouragingly, over time these effects
appear to diminish in size. One limitation with the calculation of
effect sizes for count data is the sensitivity of the data to sample
size (i.e., $\chi^2$ is upwardly biased by sample size, and $V$ is
calculated based on this value). While we used the inclusion of zero as
our boundary for "significance", the interpretation of the effects is
that most are likely small: $V$ \< .05:
`r apa_num(sum(all_effects$v <.05)/nrow(all_effects)*100)`%, \< .10:
`r apa_num(sum(all_effects$v < .10)/nrow(all_effects)*100)`%, \< .20:
`r apa_num(sum(all_effects$v < .20)/nrow(all_effects)*100)`%.

@@

In this investigation, we explored the publication rates, areas, and
researchers involved in big team science publications. Over a
half-million articles were published in nearly 15,000 journals since
1970 that qualified as big team science articles with at least ten
authors and ten different affiliations. The areas of publication aligned
to cancer and genetics research in medicine and oncology for medical and
life sciences, physics and chemistry for the physical sciences, and
psychology for the social sciences. All areas of research show an
explosion in growth in the number of publications and the number of
author included on manuscripts, replicating previous investigations into
this topic area [@sinatra2015; @wuchty2007; @hunter2008].

Our investigation expands into an exploration of the researchers who
publish in big teams focusing on career length, publication metrics, and
geopolitical affiliation. The number of earlier career scholars is
increasing in publications across the years, indicating that big teams
may be more accessible to different types of individuals, not just
older, more established researchers. This result is especially
interesting given the publish-or-perish model still present in most
institutions, as it may seem that large-scale projects could be a risky
choice for non-permanent researchers. In the authors' experience, big
team projects are often quite slow to publication, incentives may be low
for non corresponding authors if institutions do not value papers
without lead authorship, and there is no guarantee for publication even
with a large group. However, with a large team the distribution of work
could imply less effort on individual non-leading members, and research
has shown that larger-team publications do receive more citations and
appear to have higher impact [@larivière2015].

The results for number of publications by big team researchers mimics
the findings from career length, with a smaller effect size. In general,
it appears that there is a decrease in the average number of
publications a researcher has when publishing on a big team science
paper over time. This result is likely attributable to the number of
early career scholars joining projects, but also may support increased
accessibility for individuals to be involved in this type of research.
Globalization, the internet, and the focus on interdisciplinary research
are potentially driving forces behind our results, but, hopefully, the
results also point to a decline in scientific gatekeeping [@siler2015;
@lu2007].

The variability in the types of researchers involved in publications
also decreased across time in most areas of science with a decrease in
variability for career length. As mention, an increase in early career
researchers and numbers of publications could explain this effect
mathematically, potentially with other social influences mentioned
above. The variability in number of publications is decreasing in the
physical sciences, mirroring the career length results, but the opposite
effect was found in the health and social sciences. We see no clear
reason why career variability would decrease while the variability in
the number of publications would increase. The effect sizes for career
length were much larger than the effects for number of publications. One
speculation is the increasing requirements for a competitive faculty
role application. Given the limited number of positions, one potential
way to distinguish their application would be a larger number
publications in their early career [@caplow2017; @kyvik2003].

The number of geopolitical entities for researcher affiliation is
increasing over time, showing the results of globalization and the
ability to connect across time zones and cultures [@xie2014]. While our
definition of big team science required at least ten different
institutional affiliations, we did not filter papers by geopolitical
region, and thus, a manuscript could rely solely on institutions within
a single country. The physical sciences did not show an increase in
diversity of regions represented, however, it could be argued that the
development of large research centers like CERN forced earlier diversity
than other sciences (i.e., because CERN specifically recruited
scientists from sponsoring nations). The life, health, and social
sciences saw an increase in the number of regions represented with the
highest increase in the social science. This result likely corresponds
with an increased interest in big team science publications in
psychology [@coles2022; @forscher2020], and the desire to diversify the
populations represented in psychological research [@henrich2010;
@newson2021].

While publications on the whole are diversifying, we did not yet find
equality in the representation for first/corresponding author spots
versus all other authors. In general, first authors appear to be less
diverse, representing more European and North American authors, while
other authors include more Asian and African authors. These effect sizes
were often small, but the inequality still persists across years, even
if they are slowly decreasing. Diverse teams are more likely to have
papers with stronger "impact" [@jones2008; @hinnant2012; @yang2022;
@freeman2015] with higher citation metrics for more diverse author
lists. The introduction of contributorship models [e.g., CRediT;
@allen2019] will hopefully continue to push these effects down, as they
highlight each individual's contribution to a manuscript.

The limitations for this research are tied to the availability and
curation of the Scopus dataset. While the number of articles analyzed
for this investigation was large, the criteria for inclusion requires
the correct entry of author affiliations, the correct author linkages
for career length and publication rates, and the correctly marked
geopolitical entity. We had planned to analyze educational levels to
determine if the number of student coauthors (i.e., non-terminal degree)
had increased over time; however, this data was mostly blank within the
Scopus archive. Scopus is a carefully curated dataset, but these
limitations must be kept in mind when interpreting the results.
Publication language diversity was not investigated, and a previous
study indicates that the majority of publications in big databases are
in English [@albarillo2014]. Certainly, publications in non-English
languages would improve the statistics on diversity in scientific
publishing - but the English language barrier likely exists regardless
of inclusion in databases [@meneghini2007; @ramírez-castañeda2020].

Big teams have the ability to provide high-impact, important research
within scientific publishing, and this report suggests a promising trend
of increasing numbers of publications that include earlier career and
more diverse scholars. These partnerships introduce new challenges to
collaboration from interpersonal conflict, infrastructure, incentives,
to international political situations [@forscher2020]. Directed studies
into ways to navigate these situations would be beneficial for policy
makers at institutions, as well as lead teams who organize and complete
these projects. The implications for retention and promotion processes
across a broad span of regions should be explored to improve diversity
with the understanding of the differential impact of incentives for
participating in big team studies.

@@

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\endgroup
```
\newpage

# (APPENDIX) Appendix {.unnumbered}
